### 2013-04-29 inst/examples/BUGS/allen.Rmd
updating uniform priors example

Load necessary libraries,

```{r libraries, cache=FALSE, message=FALSE, warning=FALSE}
library(pdgControl) # custom SDP functions
library(nonparametricbayes) # custom 
library(reshape2) # data manipulation
library(data.table) # data manipulation
library(plyr) # data manipulation
library(ggplot2) # plotting 
library(tgp) # Gaussian Processes
library(MCMCpack) # Markov Chain Monte Carlo tools 
library(R2jags) # Markov Chain Monte Carlo tools 
library(emdbook) # Markov Chain Monte Carlo tools 
library(coda) # Markov Chain Monte Carlo tools 
```

@@

```{r plotting-options}
library(knitcitations)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE)
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

@@

```{r}
  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
          meanfn="constant", bprior="b0", BTE=c(2000,20000,2),
          m0r1=FALSE, corr="exp", trace=TRUE, 
          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
```

@@

prior_curves <- ddply(posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100)
    data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
})

# 
ggplot(posteriors) + geom_histogram(aes(x=value, y=..density..), alpha=0.7) +
    geom_line(data=prior_curves, aes(x=value, y=density), col="red", lwd=2) +
    facet_wrap(~ variable, scale="free")

ggplot(posteriors, aes(value)) + stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

@@

```{r}
jags.params=c("K","logr0","logtheta","iR","iQ") # Don't need to save "x"
#jags.inits <- function(){ list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]),"iQ"=1/0.05,"iR"=1/0.1,"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123) }

jags.params=c("K","logr0","logtheta","stdQ", "stdR")
jags.inits <- function(){
  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

@@

time<-system.time(       
  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-UPrior.txt")
)         
time <- unname(time["elapsed"]);
```

@@

# posterior distributiosn
ggplot(par_posteriors, aes(value)) + stat_density(geom="path", position="identity", alpha=0.7) +
  facet_wrap(~ variable, scale="free", ncol=2)
```

```{r}
mcmc <- as.mcmc(jagsfit)
mcmcall <- mcmc[,-2]
who <- colnames(mcmcall)
who 
mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4],mcmcall[,5])
colnames(mcmcall) <- who

pardist <- mcmcall
pardist[,4] = exp(pardist[,4]) # transform model parameters back first
pardist[,5] = exp(pardist[,5])

bayes_coef <- apply(mcmcall,2,mean)
bayes_pars <- c(bayes_coef[4], bayes_coef[1], bayes_coef[5])

>>>>

@@

```{r plotting-options}
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)

library(knitcitations)

library(ggplot2) # plotting 
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```
Load necessary libraries,

```{r libraries, cache=FALSE, message=FALSE, warning=FALSE}
library(nonparametricbayes) # loads the rest as dependencies
```

@@

Estimate the Gaussian Process (nonparametric Bayesian fit)

```{r}
  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
          meanfn="constant", bprior="b0", BTE=c(2000,10000,10),
          m0r1=FALSE, corr="exp", trace=TRUE, 
          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
```

Show traces and posteriors against priors

@@

## A list of functions has been defined in "priors", 
## Here we just make sure to compute priors over the range of values returned by the posteriors
prior_curves <- ddply(posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100)
    data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
})
```

(Note that I keep thinking this should be done with `stat_function`, but cannot `facet_wrap` across different functions, see [SO/1376967](http://stackoverflow.com/questions/1376967/using-stat-function-and-facet-wrap-together-in-ggplot2-in-r)

```{r}
# Posteriors (easier to read than histograms)
ggplot(posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

@@

```{r}
jags.params=c("K","logr0","logtheta","iR","iQ") # Don't need to save "x"
# Uniform priors on standard deviation terms
jags.params=c("K","logr0","logtheta","stdQ", "stdR")
jags.inits <- function(){
  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

@@

time_jags <- system.time(       
  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-UPrior.txt")
)         
time <- unname(time["elapsed"]);
```

@@

## priors (untransformed variables)
K_prior <- function(x) dunif(x, 1, 22000)
logr_prior <- function(x) dunif(x, -4, 2)
logtheta_prior <- function(x) dunif(x, -4, 2)
stdQ_prior <- function(x) dunif(0, 100)
stdR_prior <- function(x) dunif(0, 100)

par_priors <- list(K = K_prior, logr0 = logr_prior, logtheta = logtheta_prior, 
     stdQ = stdQ_prior, stdR = stdR_prior, deviance = function(x) 0)

par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100)
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})

# posterior distributions
ggplot(par_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

```{r}
# um, cleaner if we were just be using the long form, par_posterior
mcmc <- as.mcmc(jagsfit)
mcmcall <- mcmc[,-2]
who <- colnames(mcmcall)
who 
mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4],mcmcall[,5])
colnames(mcmcall) <- who

pardist <- mcmcall
pardist[,2] = exp(pardist[,2]) # transform model parameters back first
pardist[,3] = exp(pardist[,3])

bayes_coef <- apply(mcmcall,2,mean)
bayes_pars <- c(bayes_coef[2], bayes_coef[1], bayes_coef[3])
bayes_pars
```
